What's in a Domain? Analysis of URL Features

### 2023-08-24 paper/paper.tex
Affiliation struction
Paper update

\author{John Hawkins}
\affiliation{Getting-data-science-done.com}
\affiliation{Transitional AI Research Group}

@@

Some authors have emphasized that URLs are sequences of characters and require 
feature extraction methods that respect 
this sequential nature\cite{Le2018,Vazhayil2018}. 
They have turned to developing neural network approaches
based on convolutional or recurrent layers for learning these sequential structures.
However, the ordering of key features within URLs is predominantly fixed as shown 
in Figure \ref{fig:url_structure}. The core elements of URL structure occur at 
fixed positions within the sequence, in contrast to human language
sequences which are built from flexible grammars that allow variable positioning of 
most key elements. The inherent structure of URLs has been demonstrated to be a source 
of information that can be exploited in algorithm design \cite{Shih2004}. Furthermore,
there are psychological and social elements to URL construction (such as the use of 
common brand names in subdomains for phishing attack URLS \cite{Tupsamudre2019}), 
which emphasize the utility of feature engineering techniques that
respect the inherent structure of URLs.

@@

In this work we develop a feature engineering library that maintains an ontology of 
URL features related to underlying structure of the URL. In addition we provide a 
method for visualising the importance of features across the URL structure. 
These techniques can prove useful for analysing features for problems that
consist of moving targets and require broad sets of potential features and 
efficient feature selection \cite{Basnet2012}.
We present an open source package for generating features within these
ontologocal classes and apply them to a variety of tasks. We present our results as a
cross-task evaluation of URL features. The strength of our approach lies in the 
ability to both understand how machine learning models use URLs in decision making, 
as opposed to the black-box problem with neural network
models that treat the URL as a text sequence. In addition, the cross-task evaluation 
we perform provides insight into what works within different domains. 
This is critical as many machine learning problems that rely on URLS require
models built on relatively small datasets, so that new patterns of fraud and 
behaviour can be detected rapidly.

@@

\subsection{Data}
The data used in this study was colated from a range of sources to represent a varity of
Internet resource classification problems containing URLs as the primary feature.
The ISCX Malcious URL classification dataset contains malware, spam and phishing URLS
that need to be discriminated against a set of benign URLS\cite{Mamun2016}.
The world wide web knowledge base (WebKb) 4 Universities data set contains a wide
range of university URLs categorised into multiple topic categories\cite{Craven1998}.
The Syskill \& Webert webpage ratings dataset containing webepages across 4 categories
with a human generated categorisation for determining personal preferences in webpage
content\cite{Pazzani1996}. Finally, we utilise the Kaggle DMOZ dataset that contains
a large set of topic categorised URLs\cite{dmoz}.

These 6 URL classification problems are summarised in Table \ref{tab:data}, where
show the number of records, the cardinality of the target classes, and the proportion
of the data that belongs to the most common class in the set.

@@

We generate all URL features for each of these datasets and then apply a range of machine learning
techniques to build classifiers that can be analysed for key features for each task.

@@

The commands for installing and invoking the \emph{url2features} package from the command line are shown
in Listing \ref{code:cli}. 
When execting this code it will take the tabular data inside the file \emph{input.csv} and then apply feature 
engineering on the column named \emph{URLCOL}. The features to be created are defined by the switches 
$-host$ and $-tld$ for the host and top level domain features. By default the application outputs to
STDOUT, so in this example we are redirecting the output of the program to create a new file \emph{output.csv}.

@@

The table of results indicates that internet security problems appear much more amenable to URL based 
classification. All three of these problems result in high performing classifiers. In comparison the topic
and user preference based tasks achieve far more modest performance. It is worth noting that these problems 
are hindered by both higher cardinality in the classification targets and greater variability in dataset size,
 2 of the 3 data sets are much smaller than any of the security datasets. However, the performance on the very large DMOZ
dataset strongly suggests that, in general, URL based topic classification is likely to remain a difficult
problem.

@@

We see that there is substantial variability across all classification tasks in these experiments.
The protocol and parameters sections remain mostly devoid of signal, with the exception of the Spam URL classification
task, where we see value in the URL parameter features. The domain, path and file regions are the predominant sources
of signal across these tasks. Nevertheless, the emphasis changes between them. The subdomain appears particularly
important for the phishing URL identfication, whereas the URL target file is the focus for the WebKb classification data.

@@

Our experiments demonstrated that the various structural regions play different roles across 
these tasks, exhibiting varying levels of importance. Some patterns (like the strength of global URL features)
could be attributed to broader categories of the task, while others appear task specific.
We have observed that the domain name, with its subcomponents, is a strong and consistent contributor to the
performance of models across tasks. The source of information from within the domain varies, but it remains
a rich source of information for machine learning tasks that rely on the URL as acentral feature.

>>>>

\author{John Hawkins}
\affiliation{Transitional AI Research Group\\Getting-Data-Science-Done.com}

@@

Some authors have emphasized that URLs are sequences of characters and require 
feature extraction methods that respect 
this sequential nature\cite{Le2018,Vazhayil2018}. 
They have turned to developing neural network approaches
based on convolutional or recurrent layers for learning these sequential structures.
However, the ordering of key features within URLs is predominantly fixed as shown 
in Figure \ref{fig:url_structure}. The core elements of URL structure occur at 
fixed positions within the sequence, in contrast to human language
sequences which are built from flexible grammars that allow variable positioning of 
most key elements. The inherent structure of URLs has been demonstrated to be a source 
of information that can be exploited in algorithm design \cite{Shih2004}. Furthermore,
there are psychological and social elements to URL construction (such as the use of 
common brand names in subdomains for phishing attack URLs \cite{Tupsamudre2019}), 
which emphasize the utility of feature engineering techniques that
respect the inherent structure of URLs.

@@

In this work we develop a feature engineering library that maintains an ontology of 
URL features related to underlying structure of the URL. In addition we provide a 
method for visualising the importance of features across the URL structure. 
These techniques can prove useful for analysing features for problems that
consist of moving targets and require broad sets of potential features and 
efficient feature selection \cite{Basnet2012}.
We present an open source package for generating features within these
ontologocal classes and apply them to a variety of tasks. We present our results as a
cross-task evaluation of URL features. The strength of our approach lies in the 
ability to both understand how machine learning models use URLs in decision making, 
as opposed to the black-box problem with neural network
models that treat the URL as a text sequence. In addition, the cross-task evaluation 
we perform provides insight into what works within different domains. 
This is critical as many machine learning problems that rely on URLs require
models built on relatively small datasets, so that new patterns of fraud and 
behaviour can be detected rapidly.

@@

\subsection{Data}
The data used in this study was colated from a range of sources to represent a variety of
Internet resource classification problems containing URLs as the primary feature.
The ISCX Malicious URL classification dataset contains malware, spam and phishing URLS
that need to be discriminated against a set of benign URLs\cite{Mamun2016}.
The world wide web knowledge base (WebKb) 4 Universities data set contains a wide
range of university URLs categorised into multiple topic categories\cite{Craven1998}.
The Syskill \& Webert webpage ratings dataset containing webepages across 4 categories
with a human generated categorisation for determining personal preferences in webpage
content\cite{Pazzani1996}. Finally, we utilise the Kaggle DMOZ dataset that contains
a large set of topic categorised URLs\cite{dmoz}.

These 6 URL classification problems are summarised in Table \ref{tab:data}, where we
show the number of records, the cardinality of the target classes, and the proportion
of the data that belongs to the most common class in the set.

@@

We generate all URL features for each of these datasets and then apply a range of 
machine learning
techniques to build classifiers that can be analysed for key features on each task.

@@

The commands for installing and invoking the \emph{url2features} package from the 
command line are shown in Listing \ref{code:cli}. 
When executing this code it will take the tabular data inside the file 
\emph{input.csv} and then apply feature 
engineering on the column named \emph{URLCOL}. 
The features to be created are defined by the switches 
$-host$ and $-tld$ for the host and top level domain features. 
By default the application outputs to
STDOUT, so in this example we are redirecting the output of the program to 
create a new file \emph{output.csv}.

@@

The table of results indicates that internet security problems appear much more 
amenable to URL based classification. All three of these problems result in high 
performing classifiers. By comparison the topic and user preference classification
tasks achieve far more modest performance. It is worth noting that these problems 
are hindered by both higher cardinality in the classification targets and greater 
variability in dataset size, 2 of the 3 data sets are much smaller than any of the 
security datasets. However, the performance on the very large DMOZ
dataset strongly suggests that, in general, URL based topic classification is 
likely to remain a difficult problem.

@@

We see that there is substantial variability across all classification tasks in 
these experiments.
The protocol and parameters sections remain mostly devoid of signal, with the 
exception of the Spam URL classification task, where we see value in the URL 
parameter features. The domain, path and file regions are the predominant sources
of signal across these tasks. Nevertheless, the emphasis changes between them. 
The subdomain appears particularly
important for the phishing URL identfication, whereas the URL from domain to filename 
is important for the WebKb classification data.

@@

Our experiments demonstrated that the various structural regions play different 
roles across these tasks, exhibiting varying levels of importance. 
Some patterns (like the strength of global URL features)
could be attributed to broader categories of the task, while others appear task specific.
We have observed that the domain name, with its subcomponents, is a strong and 
consistent contributor to the performance of models across tasks. The source of 
information from within the domain varies, but it remains
a rich source of information for machine learning tasks that rely on the URL as a 
central feature.
